Effect of quality-adjusted production function inputs

### 2015-04-24 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Abstraction for formulas.

<<Time Series Variables>>=
# This block of code sets up the time-series variables 
# to be used for each country.
# Adding this level of abstraction makes it easier 
# for us to change variables later.
# It also makes it easier to see what is going on 
# in each formula throughout the document.
# Economic output
y_PT <- "iGDP"
y_UK <- "iGDP"
# Portugal undajusted inputs
y_PT <- "iGDP"
K_PT_unadj <- "iKstkS.L"
L_PT_unadj <- "iL"
E_PT_unadj <- "iXpMP"
# Partugal quality-adjusted inputs
K_PT_adj <- "iKservS.L"
L_PT_adj <- "ihLest"
E_PT_adj <- "iUMP"
# UK unadjusted inputs
K_UK_unadj <- "iKstkO.WwithRD"
L_UK_unadj <- "iL"
E_UK_unadj <- "iXp"
# UK quality-adjusted inputs
K_UK_adj <- "iKservO.WwithRD"
L_UK_adj <- "ihLest"
E_UK_adj <- "iU"
# Time variables
time_PT <- "iYear"
time_UK <- "iYear"
@

<<Formulas, echo=FALSE>>=
# Set up the formulas for each type of analysis.
# To do this, use the substitute function.

# response <- "iGDP"
# # Portugal undajusted inputs
# capital <- "iKstkS.L"
# labor <- "iL"
# time <- "iYear"
# 
# formula <- substitute( response ~ capital + labor + time,
#                        list( response = substitute(response),
#                              capital = substitute(capital),
#                              labor = substitute(labor),
#                              time = substitute(time)
#                        )
# )
# formula
# 
# 
# 
# f_PT_unadj_woe <- substitute(y ~ k + l + t, list(y = y_PT,
#                                                  k = K_PT_unadj,
#                                                  l = L_PT_unadj,
#                                                  t = time_PT))
# y_PT
# substitute(y_PT)
# eval(y_PT)
# f_PT_unadj_woe
@

@@

\begin{table} \caption{Unadjusted variables.} 
\label{tab:unadjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal           & UK                 \\
      \midrule
      Output      & \Sexpr{y_PT}       & \Sexpr{y_UK}       \\
      Capital     & \Sexpr{K_PT_unadj} & \Sexpr{K_UK_unadj} \\
      Labor       & \Sexpr{L_PT_unadj} & \Sexpr{L_UK_unadj} \\
      Energy      & \Sexpr{E_PT_unadj} & \Sexpr{E_UK_unadj} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{Unadjusted variables.} 
\label{tab:quality_adjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal         & UK               \\
      \midrule
      Output      & \Sexpr{y_PT}     & \Sexpr{y_UK}     \\
      Capital     & \Sexpr{K_PT_adj} & \Sexpr{K_UK_adj} \\
      Labor       & \Sexpr{L_PT_adj} & \Sexpr{L_UK_adj} \\
      Energy      & \Sexpr{E_PT_adj} & \Sexpr{E_UK_adj} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<Portugal C-D undadjusted without energy>>=
ptCDunadjNoE <- cdModel(formula = iGDP ~ iKstkS.L + iL + iYear, 
                        data = dataPT, constrained = TRUE)
naturalCoef(ptCDunadjNoE)
@

<<Portugal C-D undadjusted with energy>>=
ptCDunadjWithE <- cdModel(formula = iGDP ~ iKstkS.L + iL + iXpMP + iYear, 
                          data = dataPT, constrained = TRUE)
naturalCoef(ptCDunadjWithE)
@

@@

<<Portugal C-D adjusted without energy>>=
ptCDadjNoE <- cdModel(formula = iGDP ~ iKservS.L + ihLest + iYear, 
                      data = dataPT, constrained = TRUE)
naturalCoef(ptCDadjNoE)
@

<<Portugal C-D adjusted with energy>>=
ptCDadjWithE <- cdModel(formula = iGDP ~ iKservS.L + ihLest + iUMP + iYear, 
                          data = dataPT, constrained = TRUE)
naturalCoef(ptCDadjWithE)
@

@@

<<UK C-D>>=
ukCDunadjNoE <- cdModel(formula = iGDP ~ iKstkO.WwithRD + iL + iYear, 
                        data = dataUK, constrained = TRUE)
ukCDunadjWithE <- cdModel(formula = iGDP ~ iKstkO.WwithRD + iL + iXp + iYear, 
                          data = dataUK, constrained = TRUE)
ukCDadjNoE <- cdModel(formula = iGDP ~ iKservO.WwithRD + ihLest + iYear, 
                        data = dataUK, constrained = TRUE)
ukCDadjWithE <- cdModel(formula = iGDP ~ iKservO.WwithRD + ihLest + iU + iYear, 
                          data = dataUK, constrained = TRUE)
@

@@

@@

<<PT CES-kle unadjusted, warning=FALSE, eval=evalCES>>=
# Not working at this time.
# ptCESkleunadjNoE <- cesModel(formula = iGDP ~ iKstkS.L + iL + iYear, 
#                nest = c(1,2,3), data = dataPT, constrained = TRUE)
# naturalCoef(ptCESkleunadjNoE)
ptCESkleunadjWithE <- cesModel(formula = iGDP ~ iKstkS.L + iL + iXpMP + iYear, 
                            nest = c(1,2,3), data = dataPT, constrained = TRUE)
naturalCoef(ptCESkleunadjWithE)
@

@@

<<PT CES-kle quality-adjusted, warning=FALSE, eval=evalCES>>=
# Not working at this time.
# ptCESkleadjNoE <- cesModel(formula = iGDP ~ iKservS.L + ihLest + iYear, 
#                         nest = c(1,2,3), data = dataPT, constrained = TRUE)
# naturalCoef(ptCESkleadjNoE)
ptCESkleadjWithE <- cesModel(formula = iGDP ~ iKservS.L + ihLest + iUMP + iYear, 
                          nest = c(1,2,3), data = dataPT, constrained = TRUE)
naturalCoef(ptCESkleadjWithE)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
%      Without energy  & \Sexpr{naturalCoef(ptCESkleunadjNoE)$lambda}      & \Sexpr{naturalCoef(ptCESkleadjNoE)$lambda} \\
      With energy     & \Sexpr{naturalCoef(ptCESkleunadjWithE)$lambda}    & \Sexpr{naturalCoef(ptCESkleadjWithE)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<PT CES-lek unadjusted, warning=FALSE, eval=evalCES>>=
# Not working at this time.
# ptCESlekunadjNoE <- cesModel(formula = iGDP ~ iKstkS.L + iL + iYear, 
#                nest = c(2,3,1), data = dataPT, constrained = TRUE)
# naturalCoef(ptCESlekunadjNoE)
ptCESlekunadjWithE <- cesModel(formula = iGDP ~ iKstkS.L + iL + iXpMP + iYear, 
                            nest = c(2,3,1), data = dataPT, constrained = TRUE)
naturalCoef(ptCESlekunadjWithE)
@

@@

<<PT CES-lek quality-adjusted, warning=FALSE, eval=evalCES>>=
# Not working at this time.
# ptCESlekadjNoE <- cesModel(formula = iGDP ~ iKservS.L + ihLest + iYear, 
#                         nest = c(2,3,1), data = dataPT, constrained = TRUE)
# naturalCoef(ptCESlekadjNoE)
ptCESlekadjWithE <- cesModel(formula = iGDP ~ iKservS.L + ihLest + iUMP + iYear, 
                          nest = c(2,3,1), data = dataPT, constrained = TRUE)
naturalCoef(ptCESlekadjWithE)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
%      Without energy  & \Sexpr{naturalCoef(ptCESlekunadjNoE)$lambda}      & \Sexpr{naturalCoef(ptCESlekadjNoE)$lambda} \\
      With energy     & \Sexpr{naturalCoef(ptCESlekunadjWithE)$lambda}    & \Sexpr{naturalCoef(ptCESlekadjWithE)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<PT CES-ekl unadjusted, warning=FALSE, eval=evalCES>>=
# Not working at this time.
# ptCESeklunadjNoE <- cesModel(formula = iGDP ~ iKstkS.L + iL + iYear, 
#                nest = c(3,1,2), data = dataPT, constrained = TRUE)
# naturalCoef(ptCESeklunadjNoE)
ptCESeklunadjWithE <- cesModel(formula = iGDP ~ iKstkS.L + iL + iXpMP + iYear, 
                            nest = c(3,1,2), data = dataPT, constrained = TRUE)
naturalCoef(ptCESeklunadjWithE)
@

@@

<<PT CES-ekl quality-adjusted, warning=FALSE, eval=evalCES>>=
# Not working at this time.
# ptCESekladjNoE <- cesModel(formula = iGDP ~ iKservS.L + ihLest + iYear, 
#                         nest = c(3,1,2), data = dataPT, constrained = TRUE)
# naturalCoef(ptCESekladjNoE)
ptCESekladjWithE <- cesModel(formula = iGDP ~ iKservS.L + ihLest + iUMP + iYear, 
                          nest = c(3,1,2), data = dataPT, constrained = TRUE)
naturalCoef(ptCESekladjWithE)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($ek$)$l$ nesting.} 
\label{tab:lambda_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
%      Without energy  & \Sexpr{naturalCoef(ptCESeklunadjNoE)$lambda}      & \Sexpr{naturalCoef(ptCESekladjNoE)$lambda} \\
      With energy     & \Sexpr{naturalCoef(ptCESeklunadjWithE)$lambda}    & \Sexpr{naturalCoef(ptCESekladjWithE)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

>>>>

<<Time Series Variables>>=
# This block of code sets up the time-series variables 
# to be used for each country.
# Adding this level of abstraction makes it easier 
# for us to change variables later.
# It also makes it easier to see what is going on 
# in each formula throughout the document.
# Use of substitute establishes an R "name"
# that can be a component of a formula.
# Economic output
y_PT <- substitute(iGDP)
y_UK <- substitute(iGDP)
# Portugal undajusted inputs
k_PT_unadj <- substitute(iKstkS.L)
l_PT_unadj <- substitute(iL)
e_PT_unadj <- substitute(iXpMP)
# Portugal quality-adjusted inputs
k_PT_adj <- substitute(iKservS.L)
l_PT_adj <- substitute(ihLest)
e_PT_adj <- substitute(iUMP)
# UK unadjusted inputs
k_UK_unadj <- substitute(iKstkO.WwithRD)
l_UK_unadj <- substitute(iL)
e_UK_unadj <- substitute(iXp)
# UK quality-adjusted inputs
k_UK_adj <- substitute(iKservO.WwithRD)
l_UK_adj <- substitute(ihLest)
e_UK_adj <- substitute(iU)
# Time variables
t_PT <- substitute(iYear)
t_UK <- substitute(iYear)
@

<<Formulas>>=
# Set up the formulas for each type of analysis.
f_PT_unadj_noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_PT, k = k_PT_unadj, 
                    l = l_PT_unadj, t = t_PT)))
f_PT_unadj_withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_PT, k = k_PT_unadj,
                    l = l_PT_unadj, e = e_PT_unadj, t = t_PT)))
f_PT_adj_noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_PT, k = k_PT_adj,
                    l = l_PT_adj, t = t_PT)))
f_PT_adj_withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_PT, k = k_PT_adj,
                    l = l_PT_adj, e = e_PT_adj, t = t_PT)))
f_UK_unadj_noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_UK, k = k_UK_unadj,
                    l = l_UK_unadj, t = t_UK)))
f_UK_unadj_withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_UK, k = k_UK_unadj,
                    l = l_UK_unadj, e = e_UK_unadj, t = t_UK)))
f_UK_adj_noE <- as.formula(
  substitute( y ~ k + l + t,
              list( y = y_UK, k = k_UK_adj,
                    l = l_UK_adj, t = t_UK)))
f_UK_adj_withE <- as.formula(
  substitute( y ~ k + l + e + t,
              list( y = y_UK, k = k_UK_adj,
                    l = l_UK_adj, e = e_UK_adj, t = t_UK)))
@

@@

\begin{table} \caption{Unadjusted variables.} 
\label{tab:unadjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal           & UK                 \\
      \midrule
      Output      & \Sexpr{y_PT}       & \Sexpr{y_UK}       \\
      Capital     & \Sexpr{k_PT_unadj} & \Sexpr{k_UK_unadj} \\
      Labor       & \Sexpr{l_PT_unadj} & \Sexpr{l_UK_unadj} \\
      Energy      & \Sexpr{e_PT_unadj} & \Sexpr{e_UK_unadj} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

\begin{table} \caption{Unadjusted variables.} 
\label{tab:quality_adjusted_variables} 
  \begin{center}
    \begin{tabular}{r l l} 
      \toprule
      Variable    & Portugal         & UK               \\
      \midrule
      Output      & \Sexpr{y_PT}     & \Sexpr{y_UK}     \\
      Capital     & \Sexpr{k_PT_adj} & \Sexpr{k_UK_adj} \\
      Labor       & \Sexpr{l_PT_adj} & \Sexpr{l_UK_adj} \\
      Energy      & \Sexpr{e_PT_adj} & \Sexpr{e_UK_adj} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<Portugal C-D undadjusted without energy>>=
ptCDunadjNoE <- cdModel(formula = f_PT_unadj_noE, 
                        data = dataPT, constrained = TRUE)
naturalCoef(ptCDunadjNoE)
@

<<Portugal C-D undadjusted with energy>>=
ptCDunadjWithE <- cdModel(formula = f_PT_unadj_withE, 
                          data = dataPT, constrained = TRUE)
naturalCoef(ptCDunadjWithE)
@

@@

<<Portugal C-D adjusted without energy>>=
ptCDadjNoE <- cdModel(formula = f_PT_adj_noE, 
                      data = dataPT, constrained = TRUE)
naturalCoef(ptCDadjNoE)
@

<<Portugal C-D adjusted with energy>>=
ptCDadjWithE <- cdModel(formula = f_PT_adj_withE, 
                          data = dataPT, constrained = TRUE)
naturalCoef(ptCDadjWithE)
@

@@

<<UK C-D>>=
ukCDunadjNoE <- cdModel(formula = f_UK_unadj_noE, 
                        data = dataUK, constrained = TRUE)
ukCDunadjWithE <- cdModel(formula = f_UK_unadj_withE, 
                          data = dataUK, constrained = TRUE)
ukCDadjNoE <- cdModel(formula = f_UK_adj_noE, 
                        data = dataUK, constrained = TRUE)
ukCDadjWithE <- cdModel(formula = f_UK_adj_withE, 
                          data = dataUK, constrained = TRUE)
@

@@

First, look at the CES model without energy.
Nesting is unabmiguous, so this fit does not need to be repeated
for each nesting option.

<<PT CES-kl, warning=FALSE, eval=evalCES>>=
ptCESklunadjNoE <- cesModel(formula = f_PT_unadj_noE, 
                            nest = c(1,2), data = dataPT, constrained = TRUE)
naturalCoef(ptCESklunadjNoE)
ptCESkladjNoE <- cesModel(formula = f_PT_adj_noE, 
                          nest = c(1,2), data = dataPT, constrained = TRUE)
naturalCoef(ptCESkladjNoE)
@

@@

<<PT CES-kle unadjusted, warning=FALSE, eval=evalCES>>=
ptCESkleunadjWithE <- cesModel(formula = f_PT_unadj_withE, 
                            nest = c(1,2,3), data = dataPT, constrained = TRUE)
naturalCoef(ptCESkleunadjWithE)
@

@@

<<PT CES-kle quality-adjusted, warning=FALSE, eval=evalCES>>=
ptCESkleadjWithE <- cesModel(formula = f_PT_adj_withE, 
                          nest = c(1,2,3), data = dataPT, constrained = TRUE)
naturalCoef(ptCESkleadjWithE)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($kl$)$e$ nesting.} 
\label{tab:lambda_CES-kle_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(ptCESklunadjNoE)$lambda}       & \Sexpr{naturalCoef(ptCESkladjNoE)$lambda} \\
      With energy     & \Sexpr{naturalCoef(ptCESkleunadjWithE)$lambda}    & \Sexpr{naturalCoef(ptCESkleadjWithE)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<PT CES-lek unadjusted, warning=FALSE, eval=evalCES>>=
ptCESlekunadjWithE <- cesModel(formula = f_PT_unadj_withE, 
                            nest = c(2,3,1), data = dataPT, constrained = TRUE)
naturalCoef(ptCESlekunadjWithE)
@

@@

<<PT CES-lek quality-adjusted, warning=FALSE, eval=evalCES>>=
ptCESlekadjWithE <- cesModel(formula = f_PT_adj_withE, 
                          nest = c(2,3,1), data = dataPT, constrained = TRUE)
naturalCoef(ptCESlekadjWithE)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($le$)$k$ nesting.} 
\label{tab:lambda_CES-lek_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(ptCESklunadjNoE)$lambda}       & \Sexpr{naturalCoef(ptCESkladjNoE)$lambda} \\
      With energy     & \Sexpr{naturalCoef(ptCESlekunadjWithE)$lambda}    & \Sexpr{naturalCoef(ptCESlekadjWithE)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}

@@

<<PT CES-ekl unadjusted, warning=FALSE, eval=evalCES>>=
ptCESeklunadjWithE <- cesModel(formula = f_PT_unadj_withE, 
                            nest = c(3,1,2), data = dataPT, constrained = TRUE)
naturalCoef(ptCESeklunadjWithE)
@

@@

<<PT CES-ekl quality-adjusted, warning=FALSE, eval=evalCES>>=
ptCESekladjWithE <- cesModel(formula = f_PT_adj_withE, 
                          nest = c(3,1,2), data = dataPT, constrained = TRUE)
naturalCoef(ptCESekladjWithE)
@

@@

\begin{table} \caption{Portugal Solow residual results with CES and ($ek$)$l$ nesting.} 
\label{tab:lambda_CES-ekl_PT} 
  \begin{center}
    \begin{tabular}{r c c} 
      \toprule
                      & Unadjusted     & Quality-adjusted   \\
      \midrule
      Without energy  & \Sexpr{naturalCoef(ptCESklunadjNoE)$lambda}       & \Sexpr{naturalCoef(ptCESkladjNoE)$lambda} \\
      With energy     & \Sexpr{naturalCoef(ptCESeklunadjWithE)$lambda}    & \Sexpr{naturalCoef(ptCESekladjWithE)$lambda} \\
      \bottomrule
    \end{tabular}
  \end{center}
\end{table}
